%% thesis.tex 2014/04/11
%
% Based on sample files of unknown authorship.
%
% The Current Maintainer of this work is Paul Vojta.

\documentclass[phd,12pt]{ucbthesis}
%\usepackage{biblatex}
%\usepackage{longtable}
\usepackage{amsmath}
\usepackage{hyperref}

%\usepackage{lineno,hyperref}
%\modulolinenumbers[5]
%\usepackage{color}
\usepackage[table]{xcolor}
\usepackage{graphicx}
\usepackage{adjustbox}
%\usepackage{graphics}

%\usepackage{deluxetable}
%\usepackage{natbib}
%\bibliographystyle{ieee-alphabetic}

\usepackage[numbers]{natbib}


\usepackage{amssymb}

\usepackage{rotating}

% To compile this file, run "latex thesis", then "biber thesis"
% (or "bibtex thesis", if the output from latex asks for that instead),
% and then "latex thesis" (without the quotes in each case).

% Double spacing, if you want it.  Do not use for the final copy.
% \def\dsp{\def\baselinestretch{2.0}\large\normalsize}
% \dsp

% If the Grad. Division insists that the first paragraph of a section
% be indented (like the others), then include this line:
% \usepackage{indentfirst}




\begin{document}

% Declarations for Front Matter

%\title{Materials in Planet Interiors: From Microscopic to Planetary
%Scales}
\title{Modeling of Planetary Interiors: From Microscopic to Global
Scales}
\author{Sean M Wahl}
\degreesemester{Spring}
\degreeyear{2017}
\degree{Doctor of Philosophy}
\chair{Associate Professor Burkhard Militzer}
\othermembers{ Professor Bruce A. Buffett, Professor David M. Romps, Professor Eugene
Chiang}
\numberofmembers{4}
\prevdegrees{B.S. (Massachusetts Institute of Technology) 2011} 
 \field{Earth \& Planetary Science}
\emphasis{Computational Science and Engineering}
\campus{Berkeley}


% For a masters thesis, replace the above \documentclass line with
% \documentclass[masters]{ucbthesis}
% This affects the title and approval pages, which by default calls this
% document a "dissertation", not a "thesis".

\maketitle
% Delete (or comment out) the \approvalpage line for the final version.
\approvalpage
\copyrightpage

\include{abstract}

\begin{frontmatter}

\begin{dedication}
\null\vfil
\begin{center}
To my brother, Michael.\\\vspace{12pt}

\begin{quote}
    \textit{Use what talents you possess: the woods would be silent if no
birds sang except those that sang best. - anonymous
}
\end{quote}
\end{center}
\vfil\null
\end{dedication}

% You can delete the \clearpage lines if you don't want these to start on
% separate pages.

\tableofcontents
\clearpage
\listoffigures
\clearpage
\listoftables

\begin{acknowledgements}
First of all, I want to thank my advisor, Professor Burkhard Militzer, for his
beneficiary advice and continuous support. His expertise in research, programming and
teaching have been invaluable to me. His effort, availability for discussion,
attention to details, and encouragements made it a great experience to work with him.
Likewise, Professor William Hubbard of the University of Arizona, who served as a
secondary advisor, offered a wealth of expertise that allowed me shift the focus of
my thesis work for a number of projects. I would next like to thank Professor Bruce
Buffett for his aid in the various projects I pursued, along with the remaining
committee members Eugene Chiang and David Romps.  I would like to thank Jill Banfield
for giving me the opportunity and instruction in teaching Mineralogy, the course that
was most enlightening to me as an undergrad.

I owe a great deal to my group members and visitors, Shuai Zhang, Hugh Wilson, Kevin
Driver, Fran\c{c}ois Soubiran, Tanis Leonhardi, Josh Tollefson, Felipe Gonzalez and
Shoh Tagawa for the great experience working together, as well as other EPS students,
Ian Rose, Brent Delbridge, Matthew Diamond, David Mangiante, Pamala Kaercher, Qingkai
Kong, Nick Knezek, Marissa Tremblay, Sara Be and postdoc Hiro Matsui, whom I had the
pleasure to work and teach with.
I would also like to recognize all of the past and present EPS graduate student fellows who made
and continue to make the EPS department a wonderful place to work and study.
I would like to give a special thanks to Tyler Arbour, who was a wonderful mentor and
friend within the department, even in the roughest times.

I appreciate the help of the present and past EPS staff, Margie Winn, Crysthel
Catambay, Roxanne Polk, Charley Paffenbarger, Judith Coyote, Nadine Spingola-Hutton,
Allyson Weiyong Tang, Goldie Negelev, Timothy T. Teague, Elana Sabolch, Marion Banks,
Micaelee Ellswythe, Gretchen vonDuering, Jann Michael Pagdanganan, Rachel Kowalik,
Roxanne Polk, Nadine Spingola-Hutton and John Werner for their help with all manner
of questions regarding working and studying at Berkeley.  I cannot imagine how my
past five years would be without their effort in making EPS run smoothly.

I am very grateful to the members of of the Juno Science team and Juno Interior
Working group with whom I was able to contribute to a mission of a spacecraft to
another planet, a childhood dream of mine.
I want to thank Professor Dave Stevenson of Caltech, who provided my first real introduction
to Planetary Science so many years ago, and has continued to be a source of advice
and inspiration.

My musician friends including Laurel North, Ameeta Patel, John Slaymaker and the
rest of the gang at the Starry Plough, my home away from home, for always being there
to lift my spirits with some tunes.  My best friend, Rena Katz, who made my time
living in Berkeley so wonderful.

Last but not the least, I am grateful to my family---my mother, my father, and my
brother. My first teachers and collaborators, to whom I owe my love of science, and
who are a source of unwavering love and support. \\\\

I acknowledge financial support from UC Berkeley Graduate Student Fellowship, NSF,
NASA and CIDER, and computational resources including NAS, NERSC, and UC Berkeley
EPS/Astronomy clusters.

\end{acknowledgements}

\end{frontmatter}

\pagestyle{headings}

% (Optional) \part{First Part}

\include{chap1}
\include{chap2}
\include{chap3}
\include{chap4}
\include{chap5}
\include{chap6}
\include{chap7}
\include{chap8}
\include{chap9}

% \appendix
% \chapter{More Monticello Candidates}

%\printbibliography

%\nocite{*}
%\bibliographystyle{plain}
%\bibliographystyle{ieeetr}
%\bibliography{thesis}

%\bibliography{references}
%\bibliography{references.bib}

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
